---
title: "NCD_Cardiac"
author: "Capstone_Hyderabad"
date: "Wednesday, July 05, 2017"
output: word_document
---



### Load required libraries


```{r }
library(caret)
library(rpart)
library(ROCR)
library(ineq)
library(plyr)
library(dplyr)
library(ggplot2)
library(scales)
library(stringr)
library(rattle)
library(RColorBrewer)
library(lmtest)
library(pscl)
require(caret)
library(mlbench)
library(ROSE)
library(caTools)
```



### Load dataset


```{r}

setwd("C:\\Material\\CAPStone\\Scripts")
data <- read.csv("ncd_latest_062817.csv")

source("required_user_functions.R")

d_data <- subset(data, Cardiac %in% c('No','Yes') & ALIVE...DEAD == 'ALIVE')

data_d <- subset(d_data, select = c(     "FINANCIAL.GROUP","FAMILY.HISTORY","STAPLE.FOOD","ALCOHOL","Known.H.o.Smoking","Cholestrol.High","PhysicalActivity","Obese","Diabetes","Cancer","Cardiac"))
str(data_d)


colnames(data_d)[1] <- "Fin_Group"                      
colnames(data_d)[2] <- "Family_history"                       
colnames(data_d)[3] <- "Staple_Food"                          
colnames(data_d)[4] <- "Alcohol"                              
colnames(data_d)[5] <- "Smoke"                  
colnames(data_d)[6] <- "High_Cholestrol"                      
colnames(data_d)[7] <- "Phy_Activity"                      
colnames(data_d)[8] <- "Obese"                                 
colnames(data_d)[9] <- "Diabetes"                              
colnames(data_d)[10] <- "Cancer"                                
colnames(data_d)[11] <- "Cardiac"                              


cardiac_vector <- c("Fin_Group","Family_history","Staple_Food","Alcohol","Smoke","High_Cholestrol","Phy_Activity","Obese","Diabetes","Cancer")
  

data_d$Obese <- ifelse(data_d$Obese > 30,'Y','N')
data_d$Obese <- as.factor(data_d$Obese)
str(data_d)

names(data_d)

nrow(data_d)
 
count(data_d, vars=c("Cardiac"))
data_d$Fin_Group=as.numeric(factor(data_d$Fin_Group))-1
data_d$Family_history=as.numeric(factor(data_d$Family_history))-1
data_d$Staple_Food=as.numeric(factor(data_d$Staple_Food))-1
data_d$Alcohol=as.numeric(factor(data_d$Alcohol))-1
data_d$Smoke=as.numeric(factor(data_d$Smoke))-1
data_d$High_Cholestrol=as.numeric(factor(data_d$High_Cholestrol))-1
data_d$Phy_Activity=as.numeric(factor(data_d$Phy_Activity))-1
data_d$Obese=as.numeric(factor(data_d$Obese))-1
data_d$Diabetes=as.numeric(factor(data_d$Diabetes))-1
data_d$Cancer=as.numeric(factor(data_d$Cancer))-1
data_d$Cardiac=as.numeric(factor(data_d$Cardiac))
data_d$Cardiac=data_d$Cardiac-1



for (e in 1:length(cardiac_vector)){
eda_functions("Cardiac",data_d,data_d[,"Cardiac"],data_d[,e],cardiac_vector[e])
  }

```

## un balanced data
```{r}
table(data_d$Cardiac)
```
### Using ROSE package Balanced the data
```{r}
mydata1 <- ROSE(Cardiac ~ .,data=data_d, seed = 123)$data

table(mydata1$Cardiac)

```
### Split dataset  into training dataset and testing dataset in the ratio 70 % : 30 %
```{r}

set.seed(88)
split <- sample.split(mydata1, SplitRatio = 0.70)
d_train <- subset(mydata1, split == TRUE)
d_test <- subset(mydata1, split == FALSE)

```





## CART

``` {r}

formula <- Cardiac ~ Fin_Group + Family_history + Staple_Food + Alcohol +  Smoke + High_Cholestrol + Phy_Activity + Obese + Diabetes + Cardiac
title <- "Classification Tree for Cardiac"
df <- d_train
fit<-NULL
pfit <- cart_fn(df, formula, title)

 
 fancyRpartPlot(pfit)
varImp(object=fit)
  plot(varImp(fit))

```

#### Interpretation of Variable Importance for Crdiac
###### Above graph shows the top 10 important variables.high cholestrol,Cancer and diabetes are the top influencing predictors.
###### Physical activity,smoking,Financial group and Alcohol comsumption also have significant importance.


### Performance measures using training data

``` {r}
target <- d_train[,'Cardiac']
title <- 'Training dataset'
perf_measures1 <- perf_measures_cart(d_train,pfit,target,title)

print(perf_measures1)
cat("\n OA is overall accuracy \n")


```

### Performance measures using test data

``` {r}
target <- d_test[,'Cardiac']
title <- 'Test dataset'
perf_measures2 <- perf_measures_cart(d_test,pfit,target,title)

print(perf_measures2)
cat("\n OA is overall accuracy \n")


```






### Logistic Regression

```{r}
strTarget<-"Cardiac"
formula <- Cardiac ~ Fin_Group + Family_history + 
                Staple_Food + Alcohol +  Smoke + High_Cholestrol + Phy_Activity + Obese + Diabetes + Cancer
logit=logistic(formula,d_train)

summary(logit)
```

#### Interpretation of Variable Importance for cardiac based on Coefficients table shown above

###### Smoking is the only significance predicator.Rest are not having any importance at significance level of 95%

```{r}
perf <- pR2(logit)
print(perf)
```


### Logistic Regression Performance measures using training data

``` {r}
target <- d_train[,'Cardiac']
title <- 'Training dataset'

perf_measures1 <- perf_measures_logistic(d_train,logit,target,title)
print(perf_measures1)
cat("\n OA is overall accuracy \n")


```

### Logistic Regression Performance measures using test data

``` {r}
target <- d_test[,'Cardiac']
title <- 'Test dataset'
perf_measures2 <- perf_measures_logistic(d_test,logit,target,title)

print(perf_measures2)
cat("\n OA is overall accuracy \n")

varImp(logit, scale = FALSE)

```











### End of R script


```

