---
title: "NCD_Diabetes"
author: "Capstone_Hyderabad"
date: "Wednesday, July 05, 2017"
output: word_document
---



### Load required libraries


```{r }
library(caret)
library(rpart)
library(ROCR)
library(ineq)
library(plyr)
library(dplyr)
library(ggplot2)
library(scales)
library(stringr)
library(rattle)
library(RColorBrewer)
library(lmtest)
library(pscl)
require(caret)
library(mlbench)
library(DMwR)

```

### Load dataset


```{r}

setwd("C:\\Material\\CAPStone\\Scripts")
data <- read.csv("ncd_latest_062817.csv")

source("required_user_functions.R")

d_data <- subset(data, Diabetes %in% c('No','Yes') & ALIVE...DEAD == 'ALIVE')

data_d <- subset(d_data, select = c("FINANCIAL.GROUP","FAMILY.HISTORY","STAPLE.FOOD",
                                      "ALCOHOL","Known.H.o.Smoking","Cholestrol.High","PhysicalActivity","Obese","Diabetes"))
str(data_d)

colnames(data_d)[1] <- "Fin_Group"                      
colnames(data_d)[2] <- "Family_history"                       
colnames(data_d)[3] <- "Staple_Food"                          
colnames(data_d)[4] <- "Alcohol"                              
colnames(data_d)[5] <- "Smoke"                  
colnames(data_d)[6] <- "High_Cholestrol"                      
colnames(data_d)[7] <- "Phy_Activity"                      
colnames(data_d)[8] <- "Obese"                                 
colnames(data_d)[9] <- "Diabetes"                              


diabetes_vector <- c("Fin_Group","Family_history","Staple_Food","Alcohol","Smoke","High_Cholestrol","Phy_Activity","Obese")



data_d$Obese <- ifelse(data_d$Obese > 30,'Y','N')
data_d$Obese <- as.factor(data_d$Obese)
str(data_d)

names(data_d)

nrow(data_d)
 
count(data_d, vars=c("Diabetes"))


```

### Split the data into two sets each containing 1500 records of Diabetes and Not Diabetes

``` {r}
d_d <- subset(data_d,Diabetes == 'Yes')
d_n <- subset(data_d,Diabetes == 'No')
#
NCD <- "Diabetes"
displaySplitCount(d_d$Diabetes,"1",NCD)
displaySplitCount(d_n$Diabetes,"2",NCD)


set.seed(123)

smp_size1 = min(1000,nrow(d_d))
samp_ind1 <- sample(seq_len(nrow(d_d)), size = smp_size1)
d1 <- d_d[samp_ind1,]

smp_size2 = min(1000,nrow(d_n))
samp_ind2 <- sample(seq_len(nrow(d_n)), size = smp_size2)
d2 <- d_n[samp_ind2,]

print(head(d1))
print(nrow(d1))
print(head(d2))
print(nrow(d2))


displaySplitCount(d1$Diabetes,"1",NCD)
displaySplitCount(d2$Diabetes,"2",NCD)



```

### Split dataset d1 into training dataset and testing dataset in the ratio 70 % : 30 %

``` {r}
set.seed(123)
rnd <- sort(sample(nrow(d1),nrow(d1)*.7))
rnd
d1_train <- d1[rnd,]
d1_test <- d1[-rnd,]

print(nrow(d1_train))
print(nrow(d1_test))

```

### Split dataset d2 into training dataset and testing dataset

``` {r}

d2_train <- d2[rnd,]
d2_test <- d2[-rnd,]

print(nrow(d2_train))
print(nrow(d2_test))

```


### Merge d1_train (Diabetes dataset) and d2_train (Non-Diabetes dataset) to form training dataset

``` {r}

d_train <- bind_rows(d1_train,d2_train)
d_test <- bind_rows(d1_test,d2_test)

print(nrow(d_train))
print(nrow(d_test))



```

### Check the dataset the proportion of Diabetes and Non-Diabetes

``` {r}

Yes1_train <- length(which(str_trim(d1_train$Diabetes) == 'Yes'))
No1_train <- length(which(str_trim(d1_train$Diabetes) == 'No'))

cat("\n Diabetes  Count Yes: ",Yes1_train," No: ",No1_train)

Yes2_train <- length(which(str_trim(d2_train$Diabetes) == 'Yes'))
No2_train <- length(which(str_trim(d2_train$Diabetes) == 'No'))

cat("\n Diabetes  Count Yes: ",Yes2_train," No: ",No2_train)

Yes1_test <- length(which(str_trim(d1_test$Diabetes) == 'Yes'))
No1_test  <- length(which(str_trim(d1_test$Diabetes) == 'No'))

cat("\n Diabetes  Count Yes: ",Yes1_test," No: ",No1_test)

Yes2_test <- length(which(str_trim(d2_test$Diabetes) == 'Yes'))
No2_test  <- length(which(str_trim(d2_test$Diabetes) == 'No'))

###

cat("\n Train dataset file Diabetes  Count : ",nrow(d_train))
table(d_train$Diabetes)

cat("\n Test dataset file Diabetes  Count : ",nrow(d_test))
table(d_test$Diabetes)
          
```

### Remove unwanted variables

#### The following columns are not required for Diabetes - NCD:

1) "S.NO."                                 
2) "Year"                                  
3) "ALIVE...DEAD"                          
4) "RELIGION"                              
5) "Random.Blood.Sugar"                    
6) "HBA1C"                                 
7) "Specialty.of.Treatment" 
8) "Height"                                
9) "Weight"                                


``` {r}

for (e in 1:length(diabetes_vector)){
eda_functions("Diabetes",d_train,d_train[,"Diabetes"],d_train[,e],diabetes_vector[e])
  }



## CART

``` {r}

formula <- Diabetes ~ Fin_Group + Family_history + Staple_Food + Alcohol +  Smoke + High_Cholestrol + Phy_Activity + Obese
title <- "Classification Tree for Diabetes"
df <- d_train
fit<-NULL
pfit <- cart_fn(df, formula, title)

 
 fancyRpartPlot(pfit)

 varImp(object=fit)
 plot(varImp(fit))

```

#### Interpretation of Variable Importance for diabetes

###### Above graph shows the top 10 important variables.Physical activity,smoking,high cholestrol and family history are the top influencing predictors.
###### Financial group and Alcohol comsumption also have significant importance.


### Performance measures using training data

``` {r}
target <- d_train[,'Diabetes']
title <- 'Training dataset'
perf_measures1 <- perf_measures_cart(d_train,pfit,target,title)

print(perf_measures1)
cat("\n OA is overall accuracy \n")


```

### Performance measures using test data

``` {r}
target <- d_test[,'Diabetes']
title <- 'Test dataset'
perf_measures2 <- perf_measures_cart(d_test,pfit,target,title)

print(perf_measures2)
cat("\n OA is overall accuracy \n")


```






### Logistic Regression

```{r}
strTarget<-"Diabetes"
formula <- Diabetes ~ Fin_Group + Family_history + 
                Staple_Food + Alcohol +  Smoke + High_Cholestrol + Phy_Activity + Obese
logit=logistic(formula,d_train)
summary(logit)
```

#### Interpretation of Variable Importance for diabetes based on  Coefficients table shown above

###### Physical activity,high cholestrol,family history and financial group are the top influencing predictors.
###### Smoking and Alcohol comsumption also have significant importance.significance level considered is 95%


```{r}
perf <- pR2(logit)
print(perf)

```
### Logistic Regression Performance measures using training data
``` {r}
target <- d_train[,'Diabetes']
title <- 'Training dataset'

perf_measures1 <- perf_measures_logistic(d_train,logit,target,title)
print(perf_measures1)
cat("\n OA is overall accuracy \n")


```

### Logistic Regression Performance measures using test data

``` {r}
target <- d_test[,'Diabetes']
title <- 'Test dataset'
perf_measures2 <- perf_measures_logistic(d_test,logit,target,title)

print(perf_measures2)
cat("\n OA is overall accuracy \n")

varImp(logit, scale = FALSE)

```

### End of R script


```